The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-11 rational.tex
intro.
_

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructions with Negatives and Fractions}

@@

LL, GoI

Int Construction

QC

Parallel computing with constraints and backtracking -- breaking
dependancies,

@@

-- WAVEFRONT START --

But the interesting solution of trees is precisely that they involve
irrational numbers. Our usual notation for irrationals such as
{{sqrt(2) }} can actually be viewed as a thunk representing the
infinite computation {{1.4142..}}.  Irrationals in this sense have a
very different information theoretic interpretation from full real
numbers.  With Chaitin's $\Omega$ numbers
\cite{Chaitin:1975:TPS:321892.321894} he postulates the existence of
real numbers that have an infinite amount of information -- ones for
example are the solution to the question about the halting properties
of every Turning machine.

That introduces the natural numbers. No solution for this polynomial
over the algebraic numbers. We must extend the numbers with $\omega$
to get a solution. We reject this in this paper and prefer to stick to
algebraic numbers. The advantage is that all the isomorphisms are
valid numerically. With the above type we could subtract $b$ from both
sides to show that $0=1$ which is nonsense. We can however have
infinite types as long as they have algebraic solutions.

In our previous work we introduce recursive types and trace
operators. This is dangerous here because infinite loops allow us to
prolong paying the debt for as long as we want. The algebraic
$\lambda$-calculus~\cite{Vaux:2009:ALC:1630585.1630590} shows that
extreme care is needed when combining negative and fractional types
with the full $\lambda$-calculus (including recursion).

-- WAVEFRONT STOP --

We have contributed an elegant and natural notion of duality that
arises from the monoidal structure of sums and pairs. This notion of
duality has a crisp semantics, clear computational interpretations and
an information theoretic basis.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Constructions with Negatives and Fractions}

@@

\paragraph*{GoI}
The other popular model of computing with values traveling along wire
is GoI -- specifically Mackie's work. How are we related to this
model? Preliminary investogationb suggests that many of the no-cloning
based constructions can be simulated in {{langRevEE}} as well. If
there is a concrete connectoion, then by Curry-Howard the logical
interpretation of {{langRevEE}} should a logic much like LL with a
strong notion of resource preservation, but also with a strong
computational basis.

\paragraph*{Int Construction}
How are we preserving two compact closed structures -- what does this
imply about the Int construction.

\paragraph*{QC} 
Are constructions are very similar to those developed in the context
of QC. Is this how nature computes? Are the analogies to time
traveling particles and anti-particles more concrete in some way?

%% Parallel computing with constraints and backtracking -- breaking
%% dependancies,

@@

\begin{itemize}
\item Maybe the problem is that there is cancellation in the second
  case and cancellation with recursive types is problematic. However
  not all cancellation is. So what gives?

\item One idea is that irrationals correspond to an infinite amount of
  computational work. If what we are cancelling is not the same
  infinity, then things go wrong.

\item Chaitin has already shown that reals have inifite amount of
  information which is paradoxical and problematic. This is a
  different issue from that of the rational being infinite
  computations.

\item Finally, not all meaningful types are meaningful polynomials --
  ex. nat. So this is not the whole story.

%% -- WAVEFRONT START --

%% But the interesting solution of trees is precisely that they involve
%% irrational numbers. Our usual notation for irrationals such as
%% {{sqrt(2) }} can actually be viewed as a thunk representing the
%% infinite computation {{1.4142..}}.  Irrationals in this sense have a
%% very different information theoretic interpretation from full real
%% numbers.  With Chaitin's $\Omega$ numbers
%% \cite{Chaitin:1975:TPS:321892.321894} he postulates the existence of
%% real numbers that have an infinite amount of information -- ones for
%% example are the solution to the question about the halting properties
%% of every Turning machine.

%% That introduces the natural numbers. No solution for this polynomial
%% over the algebraic numbers. We must extend the numbers with $\omega$
%% to get a solution. We reject this in this paper and prefer to stick to
%% algebraic numbers. The advantage is that all the isomorphisms are
%% valid numerically. With the above type we could subtract $b$ from both
%% sides to show that $0=1$ which is nonsense. We can however have
%% infinite types as long as they have algebraic solutions.

%% In our previous work we introduce recursive types and trace
%% operators. This is dangerous here because infinite loops allow us to
%% prolong paying the debt for as long as we want. The algebraic
%% $\lambda$-calculus~\cite{Vaux:2009:ALC:1630585.1630590} shows that
%% extreme care is needed when combining negative and fractional types
%% with the full $\lambda$-calculus (including recursion).

%% -- WAVEFRONT STOP --

\paragraph*{Summary}
We have contributed an elegant and natural notion of duality that
arises from the monoidal structure of sums and pairs. This notion of
duality has a crisp semantics, clear computational interpretations and
an information theoretic basis.
